---
title: "LeadPiping"
output: github_document
---

```{r}
# Libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rgdal) # Read shape file 
library(broom) # 'fortify' the data to get a dataframe format required by ggplot2
library(viridis)
library(ggcorrplot) # Visualization of a correlation matrix
library(mgcv)
library(ggpubr)
library(cvTools)
library(stargazer) # Output regression estimate
```

## Import data
```{r setup, include=FALSE}
LeadPipes <- read_csv("LeadPipes.csv")
LeadPipes = LeadPipes %>%
  mutate(StreetPostcode = as.factor(StreetPostcode),
         WOAName = as.factor(WOAName),
         Region = as.factor(Region),
         # Convert the absence/presence of lead contamination into a numerical 0/1 variable based on LeadPresence
         LeadContaminated = ifelse(LeadPresence > 0, 1, 0))
```

## Add lat and long to the dataframe LeadPipes
```{r}
ukgrid  <- "+init=epsg:27700"
latlong <- "+init=epsg:4326"

# Create a SpatialPoints object using the sp package
sp <- SpatialPoints(LeadPipes[, c("Easting", "Northing")], proj4string = CRS(ukgrid))

# Transform the british national grid coordinates in sp to geographic coordinates using spTransform()
spt <- spTransform(sp, latlong)

# Add the longitude and latitude variables to the data frame LeadPipes
LeadPipes[, c("long", "lat")] <- coordinates(spt)
LeadPipes
```


# Plot Scotland map for visualization of lead contamination
```{r}
# Download the shapefile for Scotland from NHS
download.file("http://sedsh127.sedsh.gov.uk/Atom_data/ScotGov/ZippedShapefiles/SG_NHS_HealthBoards_2019.zip" , destfile="DATA/SG_NHS_HealthBoards_2019.zip")

# Unzip this file
system("unzip DATA/SG_NHS_HealthBoards_2019.zip")

# Read shape file for Scotland map
scotland <- readOGR(dsn = "DATA/SG_NHS_HealthBoards_2019", 
                     layer = "SG_NHS_HealthBoards_2019")

# 'fortify' the data to get a dataframe format required by ggplot2
scotland_tidy <- tidy(scotland)

# Plot LeadMeasurement and LeadPresence on the same map
ggplot() +
  geom_polygon(data=scotland_tidy, aes(x = long, y = lat, group = group), color = "black", size = 0.1, fill = "lightgrey") +
  geom_point(data=LeadPipes, aes(x = Easting, y = Northing, alpha = 0.5, color = LeadMeasurement, size = LeadPresence )) +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  coord_equal() +
  theme_void() +
  guides(alpha = FALSE) +
  labs(color = expression(paste("Average lead concentration (in ",mu,"g per litre)")), size = expression(paste("Total number of samples \n (lead concentration >= 1", mu, "g per litre)")))
```


# Plot correlation matrix
```{r}
# Select interested covariates
cor.LeadPipes = LeadPipes %>%
  select(-c(CensusHouseholdCount2011, LeadMeasurement, MedianHouseAge, BuildYearRounded, TotalPoints, LeadProportion, long, lat))

corr <- round(cor(cor.LeadPipes[sapply(cor.LeadPipes, is.numeric)]), 1)
ggcorrplot(corr, type = "lower", lab = TRUE)
```


## Explanatory analysis

# numeric variables

```{r}
ggplot(LeadPipes, aes(x = factor(UrbanRural), y = LeadPresence)) +
  geom_boxplot() +
  xlab("The Urban/Rural location class of location") +
  ylab("The number of samples detected lead")
```



```{r}
ggplot(LeadPipes, aes(x = factor(ReplacedPipes), y = LeadPresence, color = factor(ReplacedPipes))) +
 geom_boxplot() +
  guides(color = FALSE) +
 xlab("Indicator for the replacement of pipes \n (0 = not replaced, 1 = replaced)") +
 ylab("The number of samples detected lead")
```

```{r}
p1 = ggplot(LeadPipes, aes(x = Phosphorus, y = LeadMeasurement)) +
 geom_point(alpha = 0.15, color = "blue") +
 xlab("Amount of orthophosphate dosing") +
 ylab(expression(paste("Average measured lead concentration (",mu,"g per litre)"))) + 
  geom_hline(yintercept=1, linetype="dashed", color = "red")

p2 = ggplot(LeadPipes, aes(x = factor(LeadContaminated), y = Phosphorus, color = factor(LeadContaminated))) +
 geom_boxplot() +
  guides(color = FALSE) +
 xlab("Indicator for lead contamination \n (0 = absense, 1 = presence)") +
 ylab("Amount of orthophosphate dosing")
ggarrange(p1,p2, widths = c(2, 1))
```


```{r}
p3 = ggplot(LeadPipes, aes(x = BuiltAfter1970, y = LeadMeasurement)) +
 geom_point(alpha = 0.15, color = "blue") +
 xlab("Proportion of houses built after 1970") +
 ylab(expression(paste("Average measured lead concentration (",mu,"g per litre)"))) + 
  geom_hline(yintercept=1, linetype="dashed", color = "red")

p4 = ggplot(LeadPipes, aes(x = factor(LeadContaminated), y = BuiltAfter1970, color = factor(LeadContaminated))) +
 geom_boxplot() +
  guides(color = FALSE) +
 xlab("Indicator for lead contamination \n (0 = absense, 1 = presence)") +
 ylab("Proportion of houses built after 1970")
ggarrange(p3,p4, widths = c(2, 1))
```

```{r}
p5 = ggplot(LeadPipes, aes(x = BuildYear, y = LeadMeasurement)) +
 geom_point(alpha = 0.15, color = "blue") +
 xlab("Average building year") +
 ylab(expression(paste("Average measured lead concentration (",mu,"g per litre)"))) + 
  geom_hline(yintercept=1, linetype="dashed", color = "red") +
  geom_vline(xintercept = 1970, linetype = "dotted", 
                color = "blue", size = 0.5)

p6 = ggplot(LeadPipes, aes(x = factor(LeadContaminated), y = BuildYear, color = factor(LeadContaminated))) +
 geom_boxplot() +
  guides(color = FALSE) +
 xlab("Indicator for lead contamination \n (0 = absense, 1 = presence)") +
 ylab("Average building year")
ggarrange(p5,p6, widths = c(2, 1))
```

```{r}
ggplot(LeadPipes, aes(x = MeanHouseAge, y = LeadMeasurement)) +
 geom_point(alpha = 0.15, color = "blue") +
 xlab("Average time") +
 ylab("Average measured lead concentration(g per litre)")
```


```{r}
nrFolds <- 5
BS_1 <- numeric(nrFolds)
BS_2 <- numeric(nrFolds)
BS_3 <- numeric(nrFolds)
BS_4 <- numeric(nrFolds)
BS_5 <- numeric(nrFolds)
BS_6 <- numeric(nrFolds)
BS_7 <- numeric(nrFolds)
BS_8 <- numeric(nrFolds)


# generate array containing fold-number for each sample (row)
folds <- cvFolds(NROW(LeadPipes), K=nrFolds)
```

GLM
```{r}
# Follow a binomial dsitribution
for(i in 1:nrFolds){
  train <- LeadPipes[folds$subsets[folds$which != i], ] #Set the training set
  validation <- LeadPipes[folds$subsets[folds$which == i], ] #Set the validation set
  model_1 <- glm(cbind(LeadPresence, TotalPoints - LeadPresence) ~ BuiltAfter1970 + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  model_2 <- glm(LeadContaminated ~ BuiltAfter1970 + UrbanRural + Phosphorus+ BuildYear + MeanHouseAge, data = train, family = binomial)
  
  pred_1 <- predict(model_1, newdata=validation, type = "response")
  pred_2 <- predict(model_2, newdata = validation, type = "response")
  
  #convert p into q in the binomial model
  pred.convert1 <- 1-(1-pred_1)^validation$TotalPoints
  
  # Calculate Brier Score
  BS_1[i] <- mean((pred.convert1 - validation$LeadContaminated)^2)
  BS_2[i] <- mean((pred_2 - validation$LeadContaminated)^2)
}

# The average Brier score
mean(BS_1) #[1] 0.1902626
mean(BS_2) #[1] 0.1996254
```

spatial model with a binomial distribution
```{r}
for(i in 1:nrFolds){
  train <- LeadPipes[folds$subsets[folds$which != i], ] #Set the training set
  validation <- LeadPipes[folds$subsets[folds$which == i], ] #Set the validation set
  # WOAName RE #the best model
  model_3 <- gam(cbind(LeadPresence, TotalPoints - LeadPresence) ~ BuiltAfter1970 + s(WOAName, bs = "re") + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  # Region RE
  model_4 <- gam(cbind(LeadPresence, TotalPoints - LeadPresence) ~ BuiltAfter1970 + s(Region, bs = "re") + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  # a spatial smoother
  model_5 <- gam(cbind(LeadPresence, TotalPoints - LeadPresence) ~ BuiltAfter1970 + s(Easting, Northing, k=100) + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  
  pred_3 <- predict(model_3,newdata=validation, type = "response") 
  pred_4 <- predict(model_4,newdata=validation, type = "response") 
  pred_5 <- predict(model_5,newdata=validation, type = "response")
  
  # convert p into q
  pred.convert3 <- 1-(1-pred_3)^validation$TotalPoints
  pred.convert4 <- 1-(1-pred_4)^validation$TotalPoints
  pred.convert5 <- 1-(1-pred_5)^validation$TotalPoints
  # Calculate Brier Score
  BS_3[i] <- mean((pred.convert3 - validation$LeadContaminated)^2)
  BS_4[i] <- mean((pred.convert4 - validation$LeadContaminated)^2)
  BS_5[i] <- mean((pred.convert5 - validation$LeadContaminated)^2)
}

mean(BS_3)
mean(BS_4)
mean(BS_5)
# [1] 0.182483
# [1] 0.1918319
# [1] 0.2193152
```

spatial model with a Bernoulli distribution
```{r}
# GLMM with WOAName RE
for(i in 1:nrFolds){
  train <- LeadPipes[folds$subsets[folds$which != i], ] #Set the training set
  validation <- LeadPipes[folds$subsets[folds$which == i], ] #Set the validation set
  # WOAName RE
  model_6 <- gam(LeadContaminated ~ BuiltAfter1970 + s(WOAName, bs = "re") + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  # Region RE
  model_7 <- gam(LeadContaminated ~ BuiltAfter1970+s(Region, bs = "re") + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial)
  # a spatial smoother
  model_8 <- gam(LeadContaminated ~ BuiltAfter1970 + s(Easting, Northing, k=100) + UrbanRural + Phosphorus + MeanHouseAge, data = train, family = binomial) 
  
  pred_6 <- predict(model_6, newdata = validation, type = "response")
  pred_7 <- predict(model_7, newdata = validation, type = "response")
  pred_8 <- predict(model_8, newdata = validation, type = "response")
  # Calculate Brier Score
  BS_6[i] <- mean((pred_6 - validation$LeadContaminated)^2)
  BS_7[i] <- mean((pred_7 - validation$LeadContaminated)^2)
  BS_8[i] <- mean((pred_8 - validation$LeadContaminated)^2)
}

mean(BS_6)
mean(BS_7)
mean(BS_8)
# [1] 0.1985301
# [1] 0.200291
# [1] 0.2893984

var(BS_1)
var(BS_2) #minimum
var(BS_3)
var(BS_4)
var(BS_5)
var(BS_6)
var(BS_7)
var(BS_8)
# [1] 0.0007937225
# [1] 0.0003290356
# [1] 0.002195615
# [1] 0.0006730011
# [1] 0.003232067
# [1] 0.001144658
# [1] 0.0004424858
# [1] 0.01473555
```

plot of the average Brier score in each fold for each model
```{r}
#[1] 0.1902626
#[1] 0.1996254
# [1] 0.182483
# [1] 0.1918319
# [1] 0.2193152
# [1] 0.1985301
# [1] 0.200291
# [1] 0.2893984
Binomial.Bs <- data.frame(fold = 1:5, cbind(BS_1, BS_2, BS_3, BS_4, BS_5, BS_6, BS_7, BS_8))
Binomial.Bs1 <-  tidyr::pivot_longer(Binomial.Bs, cols = BS_1:BS_8,
                    names_to = "Models",
                    values_to = "Scores")
ggplot(Binomial.Bs1, aes(x = fold, y = Scores, color = Models)) + 
  geom_line() + 
  ylab("Brier score") +
  scale_colour_discrete(breaks=c("BS_1", "BS_2", "BS_3", "BS_4", "BS_5", "BS_6", "BS_7", "BS_8"),
                         labels=c("model 1", "model 2", "model 3", "model 4", "model 5", "model 6", "model 7", "model 8"))
```

The best Model 3
```{r}
Model3 <- gam(cbind(LeadPresence, TotalPoints - LeadPresence) ~ BuiltAfter1970 + s(WOAName, bs = "re") + UrbanRural + Phosphorus + MeanHouseAge, data = LeadPipes, family = binomial)
summary(Model3)

predpro <- predict(Model3, type = "response")
pred.convert <- 1-(1-predpro)^LeadPipes$TotalPoints

# Calculate the average Brier Score
ave.BS <- mean((pred.convert - LeadPipes$LeadContaminated)^2)
# [1] 0.1240927

beta <- coef(Model3)
round(exp(beta[0:5]),4)
  # (Intercept) BuiltAfter1970     UrbanRural     Phosphorus   MeanHouseAge 
  #       0.3187         0.3674         0.9172         0.9994         1.0069 
```

output the results
```{r}
stargazer(Model3, keep = c("Constant", "BuiltAfter1970", "UrbanRural", "Phosphorus", "MeanHouseAge"))
```

preform model checking
```{r}
qq.gam(Model3)
```

